Gastro-intestinal transmission of _Clostridioides difficile_ and antimicrobial-resistant bacteria in healthcare settings -- calibration of an agent-based model

### 2025-11-04 paper/paper.qmd
Renamed SD params

```{r fig-sd-model, include=TRUE}
#| echo: false
#| fig.cap: "System dynamics model. $S$ and $I$ are the number of susceptible and infected patients, respectively; $A$ and $D$ are the rates of admitted and discharged patients, respectively; $p$ is the community infection rate; $\\beta$ is the transmission coefficient; and $r$ is the rate of random (*de novo*) infection."
knitr::include_graphics("./images/sd_model.svg")
```

@@

A similar approach was used to generate outbreaks in the stochastic system dynamics model, with the difference that the parameters were changed at: 1) onset of outbreak; and 2) end of outbreak. The parameters that were changed were: 1) $\beta$ (transmission coefficient); and 2) $p$ (community infection rate). The parameter $r$ (rate of random infection) was kept constant throughout the simulation.

## Model inputs and outputs

Model input parameters were broadly grouped into: 1) toilet (e.g., number of times toilet used per day, level of contamination after use or cleaning); 2) admission (e.g., community infection rate, antimicrobial prescription rate); and 3) infection parameters (e.g., relative risk increase of antimicrobial therapy on infection).

@@

Rejection sampling \acr{ABC} was employed to fit the \acr{ABM} to the observed outbreak data [@sunnakerApproximateBayesianComputation2013]. Priors were extracted from existing literature; flat priors when data were unavailable. A summary of the priors used is available in @tbl-priors. The proportion of toilets that were shared between multiple patients was set to 60%.

| Name | Distribution | Parameters | Description | Reference |
|---------------|---------------|---------------|---------------|---------------|
| Baseline (pre-outbreak) antimicrobial prescription rate | Constant | 49.9% | Proportion of patients administered antimicrobials during admission | @versportenAntimicrobialConsumptionResistance2018 |
| Admission duration | Poisson | $\lambda = 4.6  \text{ days}$ |  | @levitkryankelixhauserastrangesekassedccoffeyrHCUPFactsFigures2007 |
| Proportion of toilets that are shared between \>1 patients | Constant | 60% | The remaining proportion (40%) of toilets were dedicated to one patient | N/A |
| Community infection rate | Uniform | $a = 0.01, b = 0.1$ |  | @crobachUnderstandingClostridiumDifficile2018 |
| Antimicrobial effect | Uniform | $a = 1.31, b = 1.87$ | Odds ratio for effect of antimicrobial therapy on risk of hospital-acquired CDI | @slimingsAntibioticsHospitalacquiredClostridium2014 |
| CDI incidence | Uniform | $a=2.8, b=12.1$ | Per 10,000 patient-days | @marraIncidenceOutcomesAssociated2020 |
| Mean toilet frequenting rate | Uniform | $a=0.5,b=5$ | Per day; used as \lambda parameter for Poisson distribution |  |
| Mean toilet cleaning rate | Uniform | $a=0.5, b=3$ | Per day; used as \lambda parameter for Poisson distribution |  |
| Toilet contamination effect | Uniform | $a=0.01,b=0.99$ | Contamination impact by infected/colonised patients through toilet use |  |
| Toilet cleaning effect | Uniform | $a=0.01, b=0.99$ | Impact of toilet cleaning on contamination |  |

@@

Firstly, latin hypercube sampling was used to generate $n$ samples from the parameter prior distributions, using a uniform distribution, such that $\theta_1,\theta_2,\ldots,\theta_n$. Next, each sample parameter point $\theta_i$ was used as an input to the agent-based simulation model, generating a total of $n$ simulated datasets. The number of daily cases of nosocomial \acr{CDI} were extracted from the model outputs, and converted to a time series of nosocomial \acr{CDI} per 1000 patient days for each month, generating simulated summary statistics $S_1,S_2,\ldots,S_n$. Since each simulation starts with an empty hospital that is gradually populated with patient admissions, results from the first 31 simulated days were discarded to ensure that the model had reached a steady state.

@@

Otherwise, $S_i$ was rejected. The parameter samples that generated the accepted simulations were used to produce a posterior distribution. $\epsilon$ was set such that the acceptance rate was approximately 2.5%, which was deemed to be a good compromise between the accuracy of the posterior distribution and the computational cost of running the simulations.

A two-step approach to calibration was taken. Firstly, ABC was used to calibrate the model pre-outbreak data to identify posterior distributions for the parameters that recapitulated the baseline rate. Next, the posterior distribution of the baseline rate model was used as a prior for a model that extended the simulation into the period of CDI outbreak and implementation of enhanced IPC measures (October 2004--May 2005 in the observed dataset) [@salgadoAnalysisOutbreakClostridium2009]. To simulate the emergence of an outbreak, the model changed parameters at a discrete time point corresponding to the outbreak onset as described in @sec-outbreak-generation. The priors used for ABC were selected to be compatible with increased transmission or infection rates, for example increased antimicrobial exposure and decreased toilet cleaning frequency. To achieve this, latin hypercube sampling was repeated, and any parameters that did not satisfy these requirements (e.g., the parameter governing the antimicrobial prescription rate was _lower_ during the outbreak period) were discarded. At a subsequent discrete time point, the model again changed parameters to simulate enhanced infection control measures. The prior distributions for these parameters were selected to decrease transmission and infection -- generally the reverse of the 'outbreak' parameters. Again, only samples from the latin hypercube that satisfied these requirements were used.

@@

\acr[first_use=TRUE, case=sentence]{ABC} was also used to calibrate the stochastic \acr{SD} model to the observed outbreak data, although the calibration was performed in a single step that encompassed the entire period of the observed data (i.e., pre-outbreak, outbreak, and control periods).

@@

To assess the sensitivity of the calibrated \acr{ABM} to changes in parameters, a single-parameter local sensitivity analysis was performed [@railsback232SensitivityAnalysis2019]. Four parameters that reflected modifiable infection control practices were chosen for the sensitivity analysis: 1) antimicrobial prescription rates; 2) the mean delay between patient infection and transfer to a side room; 3) toilet cleaning rate; and 4) toilet cleaning effect. To conduct the sensitivity analysis, all the parameters except the one of interest were fixed at the posterior mean values, determined using the rejection \acr{ABC} Next, the parameter of interest was varied through a range of values, and used as an input to the simulation model at each value. The output metric was the number of cases of hospital-acquired infection within the outbreak period. For each parameter value, the simulation was run with `r sens_n_seeds` different random number generator seeds to account for stochasticity in the model -- the output metric was the mean number of cases across the `r sens_n_seeds` runs.

@@

```{r fig-posterior-densities, include=TRUE}
#| eval: true
#| echo: false
#| fig.width: 18
#| fig.height: 18
#| warning: false
#| fig-cap: "Changes in parameters during outbreak and enhanced infection control periods. Density plots show the posterior distributions of key simulation parameters after calibration using approximate Bayesian computation."
general_params <- abc_params_outbreak_control$unadj.values %>% 
  as_tibble %>% 
  dplyr::select(-`outbreak-start`, -`outbreak-end`, -`control-end`,
                -`antibiotic-effect`,
                -`random-colonisation`) %>% 
  dplyr::select(!contains("toilet")) %>% 
  pivot_longer(cols=everything()) %>%
  mutate(type = case_when(
    str_starts(name, "o-") ~ "Outbreak",
    str_starts(name, "c-") ~ "Control",
    .default = "Baseline")) %>%
  mutate(name = str_remove(name, "^o-|^c-")) %>%
  mutate(name = str_replace_all(name, "-", " ")) %>%
  mutate(name = str_replace_all(name, "colonisation", "infection")) %>%
  mutate(name = str_to_title(name)) %>% 
  # change type to ordered factor
  mutate(type = factor(type, levels = c("Baseline", "Outbreak", "Control"))) %>%
  ggplot(aes(x = value, ..scaled.., fill = type, linetype = type)) +
  geom_density(alpha = .3) +
  facet_wrap(~name, scales = "free_x") +
  theme_minimal() +
  # remove y axis
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank()) +
  xlab("Parameter value") + 
  theme(legend.title = element_blank(),
        text=element_text(size=24)) + 
  # manually set colors
  scale_fill_manual(values = c("Baseline" = "#66c2a5", "Outbreak" = "#fc8d62", "Control" = "#8da0cb")) +
  scale_linetype_manual(values = c("Baseline" = "solid", "Outbreak" = "dashed", "Control" = "dotted")) +
  # remove gridlines
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

toilet_params <- abc_params_outbreak_control$unadj.values %>% 
  as_tibble %>% 
  dplyr::select(-`outbreak-start`, -`outbreak-end`, -`control-end`,
                -`antibiotic-effect`,
                -`random-colonisation`) %>% 
  dplyr::select(contains("toilet")) %>% 
  pivot_longer(cols=everything()) %>%
  mutate(type = case_when(
    str_starts(name, "o-") ~ "Outbreak",
    str_starts(name, "c-") ~ "Control",
    .default = "Baseline")) %>%
  mutate(name = str_remove(name, "^o-|^c-")) %>%
  mutate(name = str_replace_all(name, "-", " ")) %>%
  mutate(name = str_to_title(name)) %>% 
  # change type to ordered factor
  mutate(type = factor(type, levels = c("Baseline", "Outbreak", "Control"))) %>%
  ggplot(aes(x = value, ..scaled.., fill = type, linetype = type)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~name, scales = "free_x") +
  theme_minimal() + 
  # remove y axis
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank()) +
  xlab("Parameter value") + 
  theme(legend.title = element_blank(),
        text=element_text(size=24)) + 
  # manually set colors
  scale_fill_manual(values = c("Baseline" = "#66c2a5", "Outbreak" = "#fc8d62", "Control" = "#8da0cb")) +
  scale_linetype_manual(values = c("Baseline" = "solid", "Outbreak" = "dashed", "Control" = "dotted")) +
  # remove gridlines
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

@@

```{r run-stochastic-sd-model, include=FALSE, cache=TRUE, cache.lazy=FALSE}
set.seed(42)
randomLHS_sd <- randomLHS(1000000, 8)
sd_param_grid <- data.frame(init_population = rep(610, 1000),
                            A = 68,
                            r = qunif(randomLHS_sd[, 3], 0.001, 0.01),
                            baseline_beta = qunif(randomLHS_sd[, 4], 0.0001, 0.01),
                            baseline_p = qunif(randomLHS_sd[, 5], 0.001, 0.1),
                            outbreak_beta = qunif(randomLHS_sd[, 6], 0.0001, 0.01),
                            outbreak_p = qunif(randomLHS_sd[, 7], 0.001, 0.1),
                            outbreak_start = rep(1042, 1000),
                            outbreak_end = rep(1234, 1000),
                            n_days = rep(2585, 1000))

sd_param_grid <- sd_param_grid %>% 
  filter(baseline_beta < outbreak_beta & baseline_p < outbreak_p) %>% 
  slice_sample(n = 10000) %>% 
  mutate(param_row = row_number())
n_seeds <- 2
sd_sims <- 
  pmap_dfr(sd_param_grid,
        \(init_population, A, r, baseline_beta, baseline_p, outbreak_beta, outbreak_p, outbreak_start, outbreak_end, n_days, param_row) {
          map(1:n_seeds,
              \(seed) {
                set.seed(seed)
                sd_stochastic(init_population, A, r, baseline_beta, baseline_p, outbreak_beta, outbreak_p, outbreak_start, outbreak_end, n_days) %>% 
                  mutate(`random-seed` = seed,
                         param_row = param_row)
              }) %>% 
            bind_rows()
        }, .progress = T)

@@

posterior_predictive_sim_sd <- 
  pmap_dfr(posterior_check_params,
        \(init_population, A, r, baseline_beta, baseline_p, outbreak_beta, outbreak_p, outbreak_start, outbreak_end, n_days, seed) {
            set.seed(seed)
            sd_stochastic(init_population, A, r, baseline_beta, baseline_p, outbreak_beta, outbreak_p, outbreak_start, outbreak_end, n_days) %>% 
              mutate(`random-seed` = seed) %>% 
              bind_rows()
        }, .progress = T)

@@

```{r fig-abc-sd-posteriors, include=TRUE}
#| echo: false
#| fig.cap: "Posterior distributions of parameters obtained from approximate Bayesian computation using summary statistics generated from the stochastic deterministic model. The red line shows the prior distribution; the blue line shows the posterior distribution."

sd_posterior_params_abc <- abc_sd$unadj.values
colnames(sd_posterior_params_abc) <- names(dplyr::select(sd_param_grid, -param_row))
sd_posterior_params_abc %>% 
  as.data.frame() %>% 
  dplyr::select(r, baseline_beta, baseline_p, outbreak_beta, outbreak_p) %>% 
  pivot_longer(cols = everything(), names_to = "parameter", values_to = "value") %>% 
  mutate(parameter = factor(parameter, 
                            levels = c("r", "baseline_beta", "outbreak_beta", "baseline_p", "outbreak_p"))) %>% 
  ggplot(aes(x = value)) +
  geom_density(fill = "blue", alpha = 0.5) +
  facet_wrap(~ parameter, scales = "free")

@@

_The discussion needs to explore the different modelling approaches used (SD vs ABM) and why we eventually settled on the ABM as the main model. This is sort of captured in the results, but maybe it needs to be here as well?_

@@

>>>>

```{r fig-sd-model, include=TRUE}
#| echo: false
#| fig.cap: "System dynamics model. $S$ and $I$ are the number of susceptible and infected patients, respectively; $A$ and $D$ are the rates of admitted and discharged patients, respectively; $\\lambda$ is the community infection rate; $\\beta$ is the transmission coefficient; and $r$ is the rate of random (*de novo*) infection."
knitr::include_graphics("./images/sd_model.svg")
```

@@

A similar approach was used to generate outbreaks in the stochastic system dynamics model. To align with the agent-based model, parameters were changed at four discrete time points: 1) onset of outbreak; 2) start of enhanced IPC measures; 3) end of outbreak; and 4) end of enhanced \acr{IPC} measures. Three parameter regimes were used: baseline (pre- and post-outbreak), outbreak (pre-control), and control (during enhanced IPC). During the control period only the transmission coefficient $\\beta$ changed; the community infection rate $\\lambda$ was unaffected by control measures (it only differed between baseline and outbreak). The parameter $r$ (rate of random infection) was kept constant throughout the simulation.

@@

Rejection sampling \acr{ABC} was employed to fit the \acr{ABM} to the observed outbreak data [@sunnakerApproximateBayesianComputation2013]. Model input parameters were broadly grouped into: 1) toilet (e.g., number of times toilet used per day, level of contamination after use or cleaning); 2) admission (e.g., community infection rate, antimicrobial prescription rate); and 3) infection parameters (e.g., relative risk increase of antimicrobial therapy on infection).

Priors were extracted from existing literature; flat priors when data were unavailable. A summary of the priors used is available in @tbl-priors. The hospital design was configured based on UK standards for hospital inpatient facilities [@departmentofhealthHealthBuildingNote2013]. This consisted of 600 beds in 24 wards, of which 376 (63%) were single-occupancy rooms. Each ward had a median capacity of 24 patients, and a median of two multi-occupancy bay (each of which had four patients).

| Name | Distribution | Parameters | Description | Reference |
|---------------|---------------|---------------|---------------|---------------|
| Baseline (pre-outbreak) antimicrobial prescription rate | Constant | 49.9% | Proportion of patients administered antimicrobials during admission | @versportenAntimicrobialConsumptionResistance2018 |
| Admission duration | Poisson | $\lambda = 4.6  \text{ days}$ |  | @levitkryankelixhauserastrangesekassedccoffeyrHCUPFactsFigures2007 |
| Community infection rate | Uniform | $a = 0.01, b = 0.1$ |  | @crobachUnderstandingClostridiumDifficile2018 |
| Antimicrobial effect | Uniform | $a = 1.31, b = 1.87$ | Odds ratio for effect of antimicrobial therapy on risk of hospital-acquired CDI | @slimingsAntibioticsHospitalacquiredClostridium2014 |
| CDI incidence | Uniform | $a=2.8, b=12.1$ | Per 10,000 patient-days | @marraIncidenceOutcomesAssociated2020 |
| Mean toilet frequenting rate | Uniform | $a=0.5,b=5$ | Per day; used as \lambda parameter for Poisson distribution |  |
| Mean toilet cleaning rate | Uniform | $a=0.5, b=3$ | Per day; used as \lambda parameter for Poisson distribution |  |
| Toilet contamination effect | Uniform | $a=0.01,b=0.99$ | Contamination impact by infected/colonised patients through toilet use |  |
| Toilet cleaning effect | Uniform | $a=0.01, b=0.99$ | Impact of toilet cleaning on contamination |  |

@@

Firstly, $n$ independent random samples were drawn from the parameter prior distributions giving $\theta_1,\theta_2,\ldots,\theta_n$. Next, each sample parameter point $\theta_i$ was used as an input to the agent-based simulation model, generating a total of $n$ simulated datasets. The number of daily cases of nosocomial \acr{CDI} were extracted from the model outputs, and converted to a time series of nosocomial \acr{CDI} per 1000 patient days for each month, generating simulated summary statistics $S_1,S_2,\ldots,S_n$. Since each simulation starts with an empty hospital that is gradually populated with patient admissions, results from the first 31 simulated days were discarded to ensure that the model had reached a steady state.

@@

Otherwise, $\theta_i$ was rejected. The parameter samples that generated the accepted simulations were used to approximate a posterior distribution. $\epsilon$ was set such that the acceptance rate was approximately 2.5%, which was deemed to be a good compromise between the accuracy of the posterior distribution and the computational cost of running the simulations.

A two-step approach to calibration was taken. Firstly, ABC was used to calibrate the model pre-outbreak data to identify posterior distributions for the parameters that recapitulated the baseline rate. Next, the posterior distribution of the baseline rate model was used as a prior for a model that extended the simulation into the period of CDI outbreak and implementation of enhanced IPC measures (October 2004--May 2005 in the observed dataset) [@salgadoAnalysisOutbreakClostridium2009]. To simulate the emergence of an outbreak, the model changed parameters at a discrete time point corresponding to the outbreak onset as described in @sec-outbreak-generation. The priors used for ABC to calibrate the parameters during the outbreak period were selected to be compatible with increased transmission or infection rates, for example increased antimicrobial exposure and decreased toilet cleaning frequency, when compared to the pre-outbreak period. To achieve this, the posterior distribution of the pre-outbreak calibration was used to inform the outbreak priors. For example, the lower bound of the antimicrobial prescription rate parameter outbreak prior was set to the mean of the pre-outbreak posterior. At a subsequent discrete time point, the model again changed parameters to simulate enhanced infection control measures. The prior distributions for these parameters were selected to decrease transmission and infection -- generally the reverse of the 'outbreak' parameters.

@@

A similar calibration approach using \acr{ABC} was also used to calibrate the stochastic \acr{SD} model to the observed outbreak data, although the calibration was performed in a single step that encompassed the entire period of the observed data (i.e., pre-outbreak, outbreak, and control periods).

@@

To assess the sensitivity of the calibrated \acr{ABM} to changes in infection control practice, a single-parameter local sensitivity analysis was performed [@railsback232SensitivityAnalysis2019]. Four parameters that reflected modifiable infection control practices were chosen for the sensitivity analysis: 1) antimicrobial prescription rates; 2) the mean delay between patient infection and transfer to a side room; 3) toilet cleaning rate; and 4) toilet cleaning effect. To conduct the sensitivity analysis, all the parameters except the one of interest were fixed at the posterior mean values, determined using the rejection \acr{ABC} Next, the parameter of interest was varied through a range of values, and used as an input to the simulation model at each value. The output metric was the number of cases of hospital-acquired infection within the outbreak period. For each parameter value, the simulation was run with `r sens_n_seeds` different random number generator seeds to account for stochasticity in the model -- the output metric was the mean number of cases across the `r sens_n_seeds` runs.

@@

```{r fig-posterior-densities, include=TRUE}
#| eval: true
#| echo: false
#| fig.width: 18
#| fig.height: 18
#| warning: false
#| fig-cap: "Changes in parameters during outbreak and enhanced infection control periods. Density plots show the posterior distributions of key simulation parameters after calibration using approximate Bayesian computation."
general_params <- abc_params_outbreak_control$unadj.values %>% 
  as_tibble %>% 
  dplyr::select(-`outbreak-start`, -`outbreak-end`, -`control-end`,
                -`antibiotic-effect`,
                -`random-colonisation`) %>% 
  dplyr::select(!contains("toilet")) %>% 
  pivot_longer(cols=everything()) %>%
  mutate(type = case_when(
    str_starts(name, "o-") ~ "Outbreak",
    str_starts(name, "c-") ~ "Control",
    .default = "Baseline")) %>%
  mutate(name = str_remove(name, "^o-|^c-")) %>%
  mutate(name = str_replace_all(name, "-", " ")) %>%
  mutate(name = str_replace_all(name, "colonisation", "infection")) %>%
  mutate(name = str_to_title(name)) %>% 
  # change type to ordered factor
  mutate(type = factor(type, levels = c("Baseline", "Outbreak", "Control"))) %>%
  ggplot(aes(x = value, ..scaled.., fill = type, linetype = type)) +
  geom_density(alpha = .3) +
  facet_wrap(~name, scales = "free_x") +
  theme_minimal() +
  # remove y axis
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank()) +
  xlab("Parameter value") + 
  theme(legend.title = element_blank(),
        text=element_text(size=24)) + 
  # manually set colors
  scale_fill_manual(values = c("Baseline" = "#66c2a5", "Outbreak" = "#fc8d62", "Control" = "#8da0cb")) +
  scale_linetype_manual(values = c("Baseline" = "solid", "Outbreak" = "dashed", "Control" = "dotted")) +
  # remove gridlines
  scale_x_continuous(labels = scales::label_number(accuracy = 0.001)) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

toilet_params <- abc_params_outbreak_control$unadj.values %>% 
  as_tibble %>% 
  dplyr::select(-`outbreak-start`, -`outbreak-end`, -`control-end`,
                -`antibiotic-effect`,
                -`random-colonisation`) %>% 
  dplyr::select(contains("toilet")) %>% 
  pivot_longer(cols=everything()) %>%
  mutate(type = case_when(
    str_starts(name, "o-") ~ "Outbreak",
    str_starts(name, "c-") ~ "Control",
    .default = "Baseline")) %>%
  mutate(name = str_remove(name, "^o-|^c-")) %>%
  mutate(name = str_replace_all(name, "-", " ")) %>%
  mutate(name = str_to_title(name)) %>% 
  # change type to ordered factor
  mutate(type = factor(type, levels = c("Baseline", "Outbreak", "Control"))) %>%
  ggplot(aes(x = value, ..scaled.., fill = type, linetype = type)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~name, scales = "free_x") +
  theme_minimal() + 
  # remove y axis
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank()) +
  xlab("Parameter value") + 
  theme(legend.title = element_blank(),
        text=element_text(size=24)) + 
  # manually set colors
  scale_fill_manual(values = c("Baseline" = "#66c2a5", "Outbreak" = "#fc8d62", "Control" = "#8da0cb")) +
  scale_linetype_manual(values = c("Baseline" = "solid", "Outbreak" = "dashed", "Control" = "dotted")) +
  # remove gridlines
  scale_x_continuous(labels = scales::label_number(accuracy = 0.001)) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

@@

```{r run-stochastic-sd-model, include=FALSE, cache=TRUE, cache.lazy=FALSE}
set.seed(42)
# Sampling configuration: 'lhs' for Latin Hypercube (default) or 'iid' for independent random draws
sd_sampling_method <- "iid"  # change to "iid" for independent sampling
n_pool <- 10000
sd_dims <- 11
if (sd_sampling_method == "lhs") {
  randomLHS_sd <- randomLHS(n_pool, sd_dims)
} else {
  # independent uniform draws in [0,1]
  randomLHS_sd <- matrix(runif(n_pool * sd_dims), ncol = sd_dims)
}
# Set SD timepoint priors similar to variable_pre_outbreak.R (no ABM dependency)
eicm_start <- as.numeric(difftime(ymd('2004-11-15'), ymd('2002-01-01'), units = 'days'))
outbreak_start_center <- as.numeric(difftime(ymd('2004-10-01'), ymd('2002-01-01'), units = 'days'))
outbreak_end_center   <- as.numeric(difftime(ymd('2005-05-31'), ymd('2002-01-01'), units = 'days'))
control_end_center    <- outbreak_end_center + 90
sd_param_grid <- data.frame(init_population = rep(610, 1000),
                            A = 68,
                            r = qunif(randomLHS_sd[, 3], 0.001, 0.01),
                            baseline_beta = qunif(randomLHS_sd[, 4], 0.0001, 0.01),
                            baseline_lambda = qunif(randomLHS_sd[, 5], 0.001, 0.1),
                            outbreak_beta = qunif(randomLHS_sd[, 6], 0.0001, 0.01),
                            outbreak_lambda = qunif(randomLHS_sd[, 7], 0.001, 0.1),
                            control_beta = qunif(randomLHS_sd[, 8], 0.0001, 0.01),
                            outbreak_start = qunif(randomLHS_sd[, 9], outbreak_start_center - 90, outbreak_start_center + 90),
                            outbreak_end   = qunif(randomLHS_sd[, 10], outbreak_end_center - 90, outbreak_end_center + 90),
                            control_start  = rep(eicm_start, 1000),
                            control_end    = qunif(randomLHS_sd[, 11], control_end_center - 30, control_end_center + 30),
                            n_days = rep(2585, 1000))

sd_param_grid <- sd_param_grid %>% 
  # filter(baseline_beta < outbreak_beta & baseline_lambda < outbreak_lambda) %>% 
  # slice_sample(n = 10000) %>% 
  mutate(param_row = row_number())
n_seeds <- 5
sd_sims <- 
  pmap_dfr(sd_param_grid,
  \(init_population, A, r, baseline_beta, baseline_lambda, outbreak_beta, outbreak_lambda, control_beta, outbreak_start, outbreak_end, control_start, control_end, n_days, param_row) {
          map(1:n_seeds,
              \(seed) {
                set.seed(seed)
      sd_stochastic(init_population=init_population,
          A=A,
          r=r,
          baseline_beta = baseline_beta,
          baseline_lambda = baseline_lambda,
          outbreak_beta = outbreak_beta,
          outbreak_lambda = outbreak_lambda,
          control_beta = control_beta,
          outbreak_start = outbreak_start,
          control_start = control_start,
          outbreak_end = outbreak_end,
          control_end = control_end,
          n_days = n_days) %>% 
                  mutate(`random-seed` = seed,
                         param_row = param_row)
              }) %>% 
            bind_rows()
        }, .progress = T)

@@

posterior_predictive_sim_sd <- 
  pmap_dfr(posterior_check_params,
    \(init_population, A, r, baseline_beta, baseline_lambda, outbreak_beta, outbreak_lambda, control_beta, outbreak_start, outbreak_end, control_start, control_end, n_days, seed) {
      set.seed(seed)
    sd_stochastic(init_population = init_population,
      A = A,
      r = r,
      baseline_beta = baseline_beta,
      baseline_lambda = baseline_lambda,
      outbreak_beta = outbreak_beta,
      outbreak_lambda = outbreak_lambda,
      control_beta = control_beta,
      outbreak_start = outbreak_start,
      control_start = control_start,
      outbreak_end = outbreak_end,
      control_end = control_end,
      n_days = n_days) %>% 
        mutate(`random-seed` = seed) %>% 
        bind_rows()
        }, .progress = T)

@@

```{r fig-abc-sd-posteriors, include=TRUE, dev='png'}
#| echo: false
#| out.height: "400px"
#| fig.cap: "Changes in SD model parameters across baseline, outbreak, and control periods. Density plots show the posterior distributions of key parameters after calibration using approximate Bayesian computation."
param_names_sd <- c(
  "R" = "Community infection rate (_r_)",
  "Beta" = "Transmission rate (\U03B2)",
  "Lambda" = "Toilet contamination rate (\U03BB)"
)
sd_posterior_params_abc <- abc_sd$unadj.values
colnames(sd_posterior_params_abc) <- names(dplyr::select(sd_param_grid, -param_row))
sd_posterior_params_abc %>% 
  as.data.frame() %>% 
  # select only parameter columns (exclude timepoints)
  dplyr::select(r, baseline_beta, outbreak_beta, control_beta, baseline_lambda, outbreak_lambda) %>% 
  pivot_longer(cols = everything(), names_to = "name", values_to = "value") %>% 
  mutate(type = case_when(
    str_starts(name, "outbreak_") ~ "Outbreak",
    str_starts(name, "control_")  ~ "Control",
    TRUE                           ~ "Baseline"
  )) %>% 
  mutate(name = str_remove(name, "^baseline_|^outbreak_|^control_")) %>% 
  mutate(name = str_replace_all(name, "_", " ")) %>% 
  mutate(name = str_to_title(name)) %>% 
  mutate(type = factor(type, levels = c("Baseline", "Outbreak", "Control"))) %>%
  ggplot(aes(x = value, ..scaled.., fill = type, linetype = type)) +
  geom_density(alpha = .3) +
  facet_wrap(~ name, scales = "free_x", labeller = as_labeller(param_names_sd),
             ncol = 1) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.text = ggtext::element_markdown()) +
  xlab("Parameter value") + 
  theme(legend.title = element_blank(),
        legend.position = "top"
        # ,text=element_text(size=24)
        ) + 
  scale_fill_manual(values = c("Baseline" = "#66c2a5", "Outbreak" = "#fc8d62", "Control" = "#8da0cb")) +
  scale_linetype_manual(values = c("Baseline" = "solid", "Outbreak" = "dashed", "Control" = "dotted")) +
  scale_x_continuous(labels = scales::label_number(accuracy = 0.001)) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

@@

@@

# References
